---
title: "HCI"
output: github_document
---

This code reads data from netcdf files (downloaded from [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/sis-tourism-climate-suitability-indicators?tab=overview)) and makes plots for selected stations showing HCI values
under different RCP scenarios.

## Importing libraries and data

```{r}
source("common.R", local = knitr::knit_global())

library(ggplot2)
```

Read and process all data for HCI:

```{r cache=TRUE}
all_hci_data <- assembledata("hci")
```

All relevant data is in `alldata`, where column `stationid` is the index of 
the grid point in the raw imported data, `scenario` is either "historical", 
"RCP2.6", "RCP4.5" or "RCP8.5", `time_period` is either "1986-2005", "2021-2040", 
"2041-2060" or "2081-2100", `metric` is the mean, 10th percentile or 90th 
percentile, `day_cat` is either "fair", "good" or "unf" (category determined based on HCI value), `datapoint` is the count of days in `day_cat` category for a specific month and metric, 
`month` is the month for which the datapoint was calculated.

```{r}
head(all_hci_data)
```

## Plot

Plot data for all stations and scenarios:

```{r, dpi=300, fig.width=8, fig.height=4}
for (stat_id in gridpoint_indexes) {
    for (scen in scenarios) {
        p <- plotdata(stat_id, scen, all_hci_data)
        print(p)
    }
}
```

Save all the plots:

```{r, eval=FALSE}
for (stat_id in gridpoint_indexes) {
    for (scen in scenarios) {
        stat_name <- names[match(stat_id, gridpoint_indexes)]
        print(paste("HCI", stat_name, scen))

        p <- plotdata(stat_id, scen, all_hci_data)
        
        ggsave(paste("HCI_", gsub(" ", "_", stat_name), "_", scen, ".pdf", sep=""), p, width=8, height=4, units="in", path="../output", device=cairo_pdf)
        ggsave(paste("HCI_", gsub(" ", "_", stat_name), "_", scen, ".eps", sep=""), p, width=8, height=4, units="in", path="../output", device=cairo_ps)
        ggsave(paste("HCI_", gsub(" ", "_", stat_name), "_", scen, ".png", sep=""), p, width=8, height=4, units="in", path="../output", dpi=500)
    }
}
```